Reactor like TGE model

### 2019-06-24 AYSS-2019/article/ZelenyyAYSS-2019.tex
Fixed AYSS-2019

% My package
\usepackage{overpic}

\usepackage{todonotes}

@@

\author{Mikhail Zelenyi} % Write as First name Surname
\email[Corresponding author: ]{mihail.zelenyy@phystech.edu}
\affiliation{
Moscow Institute of Physics and Technology (State University) 
}
\affiliation{%
Institute for Nuclear Research of RAS
}%
\affiliation{
Space Research Institute of RAS
}
\author{Alexander Nozik}%
\email{altavir@gmail.com}
\author{Egor Stadnichuk}
\email{egrstadnichuk@yandex.ru}
\affiliation{
Moscow Institute of Physics and Technology (State University) 
}
\affiliation{%
Institute for Nuclear Research of RAS
}%

@@

\section{Introduction}
Many previous works point to the fact that electric field inside the thundercloud is high enough to accelerate electrons and produce secondary electromagnetic showers ~\cite{gurevich1992runaway, gurevich1999lightning,dwyer2003fundamental,dwyer2011low}.
The acceleration of electron is possible under two conditions:
\begin{itemize}
    \item The strength of the field is higher than so-called critical field. The value of the critical field equals energy loss of minimally ionizing electron per length unit.
    \item The initial electron energy is high enough to be close to minimally ionizing energy.
\end{itemize}
The major problem of all of these models is that that while field strength is higher than critical field, the multiplication coefficient for showers is rather small and can’t describe some observed effects~\cite{npm_dwyer, Stadnichuk}. To solve these problems in this article we would like to present a concept of model describing a set of phenomena occurring in thundercloud, which is based on a combined global simulation of electron and photon transport inside the whole volume of the thundercloud.
\section{Reactor like TGE model}
\begin{figure}
    \centering
    \begin{overpic}[scale=.5,percent]{figures/electric_streams.pdf}
    \put(15,53){\includegraphics[scale=.015]{figures/cell.pdf}}
    \put(20,69){\text{Acceleration cell}}
    \put(10,10){\includegraphics[scale=.15]{figures/draw.pdf}}
    \end{overpic}
    \caption{
     Processes occurring in the RL-TGE model: gamma quanta run local acceleration processes in different parts of the cloud with a multidirectional electric field.
    }
    \label{fig:rl}
\end{figure}
The crucial part of this work is additional account for escaping high energy gamma-rays in the acceleration process. Escaping photons with energy higher than 1 MeV could produce (via photo-effect or Compton-effect) new electrons with energy high enough to start separate acceleration process (Fig.~\ref{fig:rl} illustrates this process). These electrons are produced not in vicinity of first accelerator, but separated by free path of gamma ray which could be in the range from tens meters to several kilometers depending on the photon energy and atmosphere density (which depends on height). Thus these secondary accelerators could not be accounted by any local model. Additionally, the field map calculation shows that field directions in two points of a thundercloud separated by some distances are mostly uncorrelated, meaning that direction of new accelerator can be more or less random relative to the initial accelerator. If the size of a thundercloud is much larger than mean free path of gamma-ray, one can observe a chain reaction, where photons are produced in elementary accelerator cells with given local multiplication coefficient depending on local field strength and then create additional elementary accelerator cells in separate places (but some of photons could be absorbed without creating additional cell). Accounting for unproductive photons as well as loss of photons through borders of a thundercloud, one can get a global multiplicative coefficient. Obviously, in case global multiplication coefficient is larger than 1, one can observe an exponential rise in number of elementary accelerator cells and therefore total radiation level (gamma, infrared and neutron) and ionization. Also for coefficient slightly less than 1, one gets slow exponential decay of radiation background. The described model is not unlike chain reaction process in nuclear reactor, so it could be called reactor-like terrestrial gamma enhancement model (RL-TGE).

@@

For fast checking the potential of our idea, we considered a next simplified model:  
\begin{itemize}
	\item There is only one tuning parameter --- the local coefficient of gamma multiplication, describing state of the atmosphere (field strength and distribution as well as density);
	\item The number of produced of photons in a cell follows Poisson distribution, direction of momentum of a generated gamma is defined by the electric field direction;
	\item Electric field is chaotic: direction of field in cell ignition point is generated by spatial uniform distribution, but magnitude is a constant for the whole cloud;
	\item Energy of particle is not taken into account, propagation of gamma is simulated with exponential distribution with fixed mean free path;
	\item Cloud size (meaning the volume with the field) could be varied, but as a reference we take one kilometer cube, we do not track particles outside the field volume. 
\end{itemize}
This model is implemented as program\footnote{\url{https://bitbucket.org/mipt-npm/skysim/src/default/}}  on Kotlin\footnote{\url{https://kotlinlang.org/}} programming language. %\todo{reference to code}

@@

Our simple implementation of reactor like model shows potential possibility for explain the TGE phenomena, but it uses a mean value of parameter. 
Improved model clarifies rough approximations of simple model, namely:
\begin{itemize}
	\item For generating electric field map used fractal model developed by Dr. Iudin; ~\cite{Iudin:2018}
	\item The photons cross-section and energy-angular distribution sampling  is computed exactly, for tracking used algorithm of maximal cross-section;
	\item The spectrum of secondary particles from burned cell is simulated by GEANT4~\cite{ALLISON2016186}
\end{itemize}

>>>>

% My package
\usepackage{overpic}
\usepackage{hyperref}

@@

\author{Mikhail Zelenyi} % Write as First name Surname
\email[Corresponding author: ]{mihail.zelenyy@phystech.edu}
\affiliation{
Moscow Institute of Physics and Technology}
\affiliation{%
Institute for Nuclear Research of RAS
}%
\affiliation{
Space Research Institute of RAS
}
\author{Alexander Nozik}%
\email{altavir@gmail.com}
\author{Egor Stadnichuk}
\email{egrstadnichuk@yandex.ru}
\affiliation{
Moscow Institute of Physics and Technology}
\affiliation{%
Institute for Nuclear Research of RAS
}%

@@

\section{Introduction}
Many previous works point to the fact that electric field inside the thundercloud is high enough to accelerate electrons and produce secondary electromagnetic showers ~\cite{gurevich1992runaway, gurevich1999lightning,dwyer2003fundamental,dwyer2011low,gurevich2001kinetic, dwyer2015positron, dwyer2013properties}.
The acceleration of electron is possible under two conditions:
\begin{itemize}
    \item The strength of the field is higher than so-called critical field. The value of the critical field equals energy loss of minimally ionizing electron per length unit.
    \item The initial electron energy is high enough to be close to minimally ionizing energy.
\end{itemize}
The major problem of all of these models is that while field strength is higher than critical field, the multiplication coefficient for showers is rather small and can’t describe some observed effects~\cite{npm_dwyer, Stadnichuk}. To solve these problems in this article we would like to present a concept of model describing a set of phenomena occurring in thundercloud, which is based on a combined global simulation of electron and photon transport inside the whole volume of the thundercloud.
\section{Reactor like TGE model}
\begin{figure}
    \centering
    \begin{overpic}[scale=.5]{figures/electric_streams.pdf}
    \put(15,53){\includegraphics[scale=.015]{figures/cell.pdf}}
    \put(20,69){\text{Acceleration cell}}
    \put(10,10){\includegraphics[scale=.15]{figures/draw.pdf}}
    \end{overpic}
    \caption{
     Processes occurring in the RL-TGE model: gamma quanta run local acceleration processes in different parts of the cloud with a multidirectional electric field.
    }
    \label{fig:rl}
\end{figure}
The crucial part of this work is additional account for escaping high energy gamma-rays in the acceleration process. Escaping photons with energy higher than 1 MeV could produce (via photo-effect or Compton-effect) new electrons with energy high enough to start separate acceleration process (Fig.~\ref{fig:rl} illustrates this process). These electrons are produced not in vicinity of first accelerator, but separated by free path of gamma ray which could be in the range from tens meters to several kilometers depending on the photon energy and atmosphere density (which depends on height). Thus these secondary accelerators could not be accounted by any local model. Additionally, the field map calculation shows that field directions in two points of a thundercloud separated by some distances are mostly uncorrelated, meaning that direction of new accelerator can be more or less random relative to the initial accelerator. If the size of a thundercloud is much larger than mean free path of gamma-ray, one can observe a chain reaction, where photons are produced in elementary accelerator cells with given local multiplication coefficient depending on local field strength and then create additional elementary accelerator cells in separate places (but some of photons could be absorbed without creating additional cell). Accounting for unproductive photons as well as loss of photons through borders of a thundercloud, one can get a global multiplicative coefficient. Obviously, in case global multiplication coefficient is larger than 1, one can observe an exponential rise in number of elementary accelerator cells and therefore total radiation level (gamma, infrared and neutron) and ionization. Also for coefficient slightly less than 1, one gets slow exponential decay of radiation background. The described model is not unlike chain reaction process in nuclear reactor, so it could be called reactor-like terrestrial gamma enhancement model (RL-TGE).

@@

For fast checking the potential of our idea, we considered a next simplified model:  
\begin{itemize}
	\item There is only one tuning parameter --- the local coefficient of gamma multiplication, describing state of the atmosphere (field strength and distribution as well as density);
	\item The number of produced of photons in a cell follows Poisson distribution, direction of momentum of a generated gamma is defined by the electric field direction;
	\item Electric field is chaotic: direction of field in cell ignition point is generated by spatial uniform distribution, but magnitude is a constant for the whole cloud;
	\item Energy of particle is not taken into account, propagation of gamma is simulated with exponential distribution with fixed mean free path;
	\item Cloud size (meaning the volume with the field) could be varied, but as a reference we take one kilometer cube, we do not track particles outside the field volume. 
\end{itemize}
This model is implemented as program\footnote{\url{https://bitbucket.org/mipt-npm/skysim/src/default/}}  on Kotlin\footnote{\url{https://kotlinlang.org/}} programming language.

@@

Our simple implementation of reactor like model shows potential possibility for explain the TGE phenomena, but it uses a mean value of parameter. 
Improved model clarifies rough approximations of simple model, namely:
\begin{itemize}
	\item For generating electric field map used fractal model developed by Dr. Iudin~\cite{Iudin:2018};
	\item The photons cross-section and energy-angular distribution sampling  is computed exactly, for tracking used algorithm of maximal cross-section;
	\item The spectrum of secondary particles from burned cell is simulated by GEANT4~\cite{ALLISON2016186}.
\end{itemize}
